---
title: "Logistic Regression Presentation"
author: "Delph, Davis, Harrison"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: true
    # lib_dir: libs
    css: [appstate.css, appstate-fonts.css]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [top, right, inverse]
      # autoplay: 40000
      # countdown: 30000
---
class: inverse, center, middle

```{r label = "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE)
library(ISLR)
library(tidyverse)
```

```{r, echo = FALSE, include=FALSE}
#Reading in the data set
file = "./Data/imports-85.data"
data <- read.table(file, sep = "," , header = F , nrows = 205,
                     na.strings ="", stringsAsFactors= F)
names(data)[1] <- "Symboling" #
names(data)[2] <- "Normalized_Loss" 
names(data)[3] <- "Make" # Make of the Vehicle 
names(data)[4] <- "Fuel-Type" # Diesel or Gas 
names(data)[5] <- "Aspiration" # Standard or Turbo
names(data)[6] <- "Doors" # Number of Doors 
names(data)[7] <- "Body" # Body Style 
names(data)[8] <- "Drive" # 4WD or 2WD
names(data)[9] <- "Engine_Loc" # Engine Location 
names(data)[10] <- "Wheel_base" # continuous 
names(data)[11] <- "length" # continuous
names(data)[12] <- "width" #continuous 
names(data)[13] <- "height" #continuous
names(data)[14] <- "curb-weight" #continuous 
names(data)[15] <- "engine-type" #dohc, dohcv, l, ohc, ohcf, ohcv, rotor.
names(data)[16] <- "num-of-cylinders" 
names(data)[17] <- "engine-size" #continuous 
names(data)[18] <- "fuel-system"
names(data)[19] <- "bore" #continuous 
names(data)[20] <- "stroke" #continuous 
names(data)[21] <- "compression-ratio" #continuous
names(data)[22] <- "horsepower" #continuous 
names(data)[23] <- "peak-rpm" #continuous 
names(data)[24] <- "city-mpg" #continuous 
names(data)[25] <- "highway-mpg" #continuous 
names(data)[26] <- "price" #continuous 
```


```{r, echo = FALSE}
# Dr.A comments
cars <- read.csv("./Data/imports-85.data", 
                 col.names=c("symbolizing", "normalized_loss", 
                             "make", "fuel_type", "aspiration",
                             "num_of_doors", "body_type", 
                             "drive_wheels", "engine_location",
                             "wheel_base","length", "width", 
                             "height", "curb_weight", "engine_type",
                             "num_of_cylinders", "engine_size",
                             "fuel_system", "bore", "stroke",
                             "compression_ratio", "horsepower", 
                             "peak_rpm", "city_mpg", "highway_mpg",
                             "price"), na.strings = "?")
cars[1:6, 1:4]
cars <- na.omit(cars)
```



```{r, include= FALSE}
library(dplyr)
mod_data <- data %>% mutate(`Fuel-Type`=ifelse(`Fuel-Type`=="gas",0,1)) %>% mutate(price = as.numeric(price))
glimpse(mod_data)
```

```{r, echo = FALSE}
# Dr. A comment
library(dplyr)
cars2 <- cars %>% mutate(fuel_type = ifelse(fuel_type == "gas", 0, 1)) 
glimpse(cars2)
```


# Dataset Information

* Our dataset was the Automobile Data Set, which included 26 attributes 

* In order to achieve a binomial variable, we selected `Fuel-Type` and mutated the original character response to a 0 = gas and 1 = diesel

* For our model and prediction, we decided to use our `Fuel-Type` variable as a function of `price`. `Price` was mutated from a character to a numeric utilizing the `as_numeric()` function.

* We hypothesized that on average diesel vehicles will cost more than gas vehicles. 

---
# Fuel-Type as a function of Price:

```{r, echo = FALSE, include = FALSE}
model <- glm(`Fuel-Type`~price, family = binomial, data = mod_data)
coef(model)
```
$$\frac{\widehat{y}}{1-\widehat{y}} = -2.746 + .000038123 * Price$$
```{r}
# Dr. A comment
mod <- glm(fuel_type ~ price, family = binomial, data = cars2)
coef(mod)
```

Note: the quantity $\pi(x)=E(Y|x)$ is used to represent the conditional mean of $Y$ given $x$ when the logistic regression is the _logit transformation_.  This transformation is defined, in terms of $\pi(x)$, as:

$$g(x) = \ln \left(\frac{\pi(x)}{1 - \pi(x)}\right) = \beta_0 + \beta_1x$$
---
# Using `ggplot2` to illustrate the model

```{r, echo = FALSE, fig.align = "center", fig.width = 7, fig.height = 5}
#Dr A comments
ggplot(data = cars2, aes(x = price, y = fuel_type)) + 
  geom_jitter(width = 0, height = 0.05, alpha = 0.1) + 
  theme_bw() + 
  labs(x = "Price", 
       y = "Fuel Type (Gas and Diesel)",
       title = "Fuel Type General Linear Model",
       subtitle = "Fuel Type as a funtion of price (Gas = 0, Diesel = 1)") + 
  geom_smooth(method = "glm", se = FALSE, 
              method.args = list(family = binomial(link = "logit")),
              color = "red") 
```
It appears as if there is a slight increase in price near the end of the model. However, this does not appear to reflect our hypothesis very well. 

---

# Prediction

```{r, echo=FALSE}
library(broom)
```
```{r}
pred <- data.frame(price = c(15000, 25000, 40000))
augment(model, newdata = pred, type.predict = "response")
```

Based on this prediction, a vehicle that cost 15K, 25K, and 40K have a 10%, 14%, and 23% probability of being a diesel. 

---
## Conclusions:

Our hypothesis assumptions may be flawed due to high-end, luxury sports cars being gas rather than diesel. 

It doesn't appear that our model is a very strong predictor because: 

--

1. There is only a slight slope to our model

--

2. The predicted probabilities are rather small

--

As such, this model is not very useful, but at least we tried

---

```{r}
# Dr A comments
summary(mod)$coef
```

Note: The results suggest that the change in log-odds of `fuel_type` per one dollar increase in `price` is $`r summary(mod)$coef[2,1]`$.

Note: The estimated logistic probability of a diesel vehicle with a price of $15,000 is:

$$\pi(15000) = \frac{\exp\hat{g}(15000)}{1 + \exp\hat{g}(15000)} = \frac{\exp-2.73 + 3.8\times 10^{-5}\times15000}{1 + \exp-2.73 + 3.8\times 10^{-5}\times15000} = 0.1025973$$

```{r}
nullmod <- glm(fuel_type ~ 1, family="binomial", data = cars2)
PR2 <- 1 - logLik(mod)/logLik(nullmod)
PR2 # McFadden's R2 --- Can we get a better model?
```
 
---

```{r}
mymod <- glm(fuel_type ~ aspiration + height + bore + stroke, data = cars2, family = "binomial")
summary(mymod)
nullmod <- glm(fuel_type ~ 1, family="binomial", data = cars2)
```

---

```{r}
PR2 <- 1 - logLik(mymod)/logLik(nullmod)
PR2 # McFadden's R2 --- Can we get a better model?
```

---


```{r}
mymod2 <- glm(fuel_type ~ horsepower + price + horsepower:price
              , data = cars2, family = "binomial")
summary(mymod2)
```

---

```{r}
PR2 <- 1 - logLik(mymod2)/logLik(nullmod)
PR2 # McFadden's R2 --- Can we get a better model?
```

```{r}
mymod3 <- glm(fuel_type ~ price + engine_size + engine_size:horsepower, data = cars2, family = "binomial")
summary(mymod3)
```


---

How well are we doing?

```{r}
DG <- predict(mymod3, type ="response")
DG <- (ifelse(DG > .5, "diesel", "gas"))
p_class <- factor(DG, levels = c("diesel", "gas"))
library(caret)
confusionMatrix(p_class, cars$fuel_type)
```



