---
title: "Correlation Regression Presentation---Group#1"
author: "Davis, Delph, Harrison"
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
output: 
  ioslides_presentation:
    logo: "./PICS/UCI.png"
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
##Load file
file = "./Data/auto-mpg.data"
data <- read.table(file, sep = "" , header = F , nrows = 398,
                     na.strings ="", stringsAsFactors= F)
carData <- as.tibble(data)
##Rename Column heads
names(carData)[1] <- "mpg"
names(carData)[2] <- "cylinders"
names(carData)[3] <- "displacement"
names(carData)[4] <- "horsepower"
names(carData)[5] <- "weight"
names(carData)[6] <- "acceleration"
names(carData)[7] <- "modelYear"
names(carData)[8] <- "origin"
names(carData)[9] <- "carName"

##remove 6 data points missing horsepower data
carDataReduced <- carData
carDataReduced$horsepower <- replace(carDataReduced$horsepower, carDataReduced$horsepower == "?", NA)
carDataReduced <- na.omit(carDataReduced)
carDataReduced$horsepower <- as.numeric(carDataReduced$horsepower)
```

## Correlation Analysis 
For our dataset, we selected `Auto MPG` and conducted correlation and regression analysis on fuel efficiency, which was measured in miles per gallon (mpg). The `Auto MPG` dataset included 8 variables, 2 of which were read in as categorical. However, after some data cleaning, we were able to remove the data points missing horsepower values to develop a regression function for horsepower.

Correlations were run for the 7 eligible variables, and the highest correlations were observed as follows: 

* MPG and Displacement produced r = - .804 

* MPG and Weight produced r = - .831

## Correlation Code

```{r, echo = TRUE, results='hide'}
carData %>% 
  summarize(N=n(), r = cor(displacement, mpg, use = "pairwise.complete.obs"))
carData %>%
  summarize(N=n(), r = cor(weight, mpg, use = "pairwise.complete.obs"))
```

## Regression: Diplacement Graph 

```{r, echo=FALSE, message=FALSE}
ggplot(carData, aes(displacement, mpg))+ 
  geom_point()+ 
  geom_smooth(method = "lm", se = FALSE)+
  theme_bw()+
  labs(title = "Fuel Efficiency",
  subtitle = "MPG as a function of Displacement",
  x = "Displacement",
  y = "MPG"
  )
```

## Regression: Displacement Code

```{r, echo=TRUE, message=FALSE, eval=FALSE}
ggplot(carData, aes(displacement, mpg))+ geom_point()+ 
  geom_smooth(method = "lm", se = FALSE)+
  theme_bw()+labs(title = "Fuel Efficiency",
  subtitle = "MPG as a function of Displacement", 
  x = "Displacement",
  y = "MPG"
  )
```

## Linear Model: Displacement
```{r, echo=TRUE}
disp_model <- lm(mpg~displacement, data = carData)
```

$$\widehat{MPG} = 35.175 + -.06\cdot{displacement}$$ 


Interpretation: 


For every unit increase in displacement, fuel efficiency is expected to decrease by 0.06 mpg. 

## Regression: Weight Graph 
```{r, echo=FALSE, message=FALSE}
ggplot(carData, aes(weight, mpg))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  theme_bw()+
  labs(
    title = "Fuel Efficiency",
    subtitle = "MPG as a function of Weight",
    x = "Weight",
    y = "MPG"
  )
```

## Regression: Weight Code 
```{r, echo=TRUE,message =FALSE, eval=FALSE}
ggplot(carData, aes(weight, mpg))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  theme_bw()+
  labs(
    title = "Fuel Efficiency",
    subtitle = "MPG as a function of Weight",
    x = "Weight",
    y = "MPG"
  )
```

## Linear Model: Weight 
```{r, echo=TRUE}
wgt_model <- lm(mpg~weight, data = carData)
```
$$\widehat{MPG} = 46.317 + -.008\cdot{weight}$$


Interpretation: 


For every unit increase in weight, fuel efficiency is expected to decrease by 0.008 mpg. 

