
# Parallel Slopes

In this chapter you'll learn about the class of linear models called "parallel slopes models." These include one numeric and one categorical explanatory variable.

____________

### What if you have two groups?{-}

<iframe src="https://drive.google.com/file/d/1eNCKcB3uJ0Piv_I1Ohfu2rLW4BNJcKPv/preview" width="740" height="420"></iframe>

____________

## Fitting a parallel slopes model

We use the `lm()` function to fit linear models to data. In this case, we want to understand how the price of mariokart games sold at auction varies as a function of not only the number of wheels included in the package, but also whether the item is new or used. Obviously, it is expected that you might have to pay a premium to buy these new. But how much is that premium? Can we estimate its value after *controlling for the number of wheels*?

We will fit a parallel slopes model using `lm()`. In addition to the `data` argument, `lm()` needs to know which variables you want to include in your regression model, and how you want to include them. It accomplishes this using a `formula` argument. A simple linear regression formula looks like `y ~ x`, where `y` is the name of the response variable, and `x` is the name of the explanatory variable. Here, we will simply extend this formula to include multiple explanatory variables. A parallel slopes model has the form `y ~ x + z`, where `z` is a categorical explanatory variable, and `x` is a numerical explanatory variable.

The output from `lm()` is a model object, which when printed, will show the fitted coefficients.

_______________

### Exercise {-}

* The dataset `mariokart` is already loaded for you. Explore the data using `glimpse()` or `str()`.

```{r}
library(openintro)
data(mariokart)
glimpse(mariokart)
# Or
# str(mariokart)
# Data munging to agree with DataCamp mario_kart
mario_kart <- mariokart %>% 
  filter(total_pr < 100)
str(mario_kart)
save(mario_kart,file = "./Data/mario_kart.RData")
```


* Use `lm()` to fit a parallel slopes model for total price as a function of the number of wheels and the condition of the item. Use the argument `data` to specify the dataset you're using.

```{r}
# fit parallel slopes
lm(total_pr ~ wheels + cond, data = mario_kart)
```
_____________________

### Reasoning about two intercepts{-}

The `mariokart` data contains several other variables. The `total_pr`, `start_pr`, and `ship_pr` variables are numeric, while the `cond` and `stockPhoto` variables are categorical.

Which formula will result in a parallel slopes model?

* `total_pr ~ start_pr + ship_pr`

* `cond ~ start_pr + stock_photo`

* **`total_pr ~ ship_pr + stock_photo`**

* `total_pr ~ cond`

__________________

### Visualizing parallel slopes models{-}

<iframe src="https://drive.google.com/file/d/13GQTCHci-gQ7Hd9MUAmtjxa6t6StJLzG/preview" width="740" height="420"></iframe>

____________


## Using `geom_line()` and `augment()`

Parallel slopes models are so-named because we can visualize these models in the data space as not one line, but two parallel lines. To do this, we'll draw two things:

* a scatterplot showing the data, with color separating the points into groups

* a line for each value of the categorical variable

Our plotting strategy is to compute the fitted values, plot these, and connect the points to form a line. The `augment()` function from the `broom` package provides an easy way to add the fitted values to our data frame, and the `geom_line()` function can then use that data frame to plot the points and connect them.

Note that this approach has the added benefit of automatically coloring the lines appropriately to match the data.

You already know how to use `ggplot()` and `geom_point()` to make the scatterplot. The only twist is that now you'll pass your `augment()`-ed model as the data argument in your `ggplot()` call. When you add your `geom_line()`, instead of letting the `y` aesthetic inherit its values from the `ggplot()` call, you can set it to the `.fitted` column of the `augment()`-ed model. This has the advantage of automatically coloring the lines for you.

_______________

### Exercise {-}

The parallel slopes model `mod` relating total price to the number of wheels and condition is already in your workspace.

```{r}
mod <- lm(formula = total_pr ~ wheels + cond, data = mario_kart)
```


* `augment()` the model `mod` and explore the returned data frame using `glimpse()`. Notice the new variables that have been created.

```{r}
library(broom)
augmented_mod <- augment(mod)
glimpse(augmented_mod)
```

* Draw the scatterplot and save it as `data_space` by passing the `augment()`-ed model to `ggplot()` and using `geom_point()`.

```{r}
# scatterplot, with color
data_space <- ggplot(data = augmented_mod, 
                     aes(x = wheels, y = total_pr, 
                         color = cond)) + 
  geom_point()
```

* Use `geom_line()` once to add two parallel lines corresponding to our model.

```{r}
# single call to geom_line()
data_space + 
  geom_line(aes(x = wheels, y = .fitted)) + 
  theme_bw()
```

### An alternative approach using `geom_parallel_slopes()` from `moderndive`

```{r}
library(moderndive)
ggplot(data = mario_kart, aes(x = wheels, y = total_pr, color = cond)) +
  geom_point() +
  geom_parallel_slopes(se = FALSE) + 
  theme_bw()
```


____________

### Interpreting parallel slopes coefficients{-}


<iframe src="https://drive.google.com/file/d/1PL0tE6AdxJvQEZUWFnX2rycsvXdJbX5m/preview" width="740" height="420"></iframe>
________________

### Intercept interpretation{-}

Recall that the `cond` variable is either `new` or `used`. Here are the fitted coefficients from your model:

```{r}
lm(total_pr ~ wheels + cond, data = mario_kart)
```
Choose the correct interpretation of the coefficient on `condused`:

* For each additional wheel, the expected price of a used mariokart is $5.58 lower.

* **The expected price of a used mariokart is $5.58 less than that of a new one with the same number of wheels.**

* The expected price of a new mariokart is $5.58 less than that of a used one with the same number of wheels.

* The used mariokarts are always $5.58 cheaper.

_________________

### Common slope interpretation{-}

Recall the fitted coefficients from our model:

```{r}
lm(total_pr ~ wheels + cond, data = mario_kart)
```

Choose the correct interpretation of the slope coefficient:

* **For each additional wheel, the expected price of a mariokart increases by $7.23 regardless of whether it is new or used.**

* For each additional wheel, the expected price of a new mariokart increases by $7.23.

* The expected price of a used mariokart is $5.59 less than that of a new one with the same number of wheels.

* You should always expect to pay $42.37 for a mariokart.

_______

### Three ways to describe a model {-}

<iframe src="https://drive.google.com/file/d/1GyJEY2vUcaru--mGOCoZa9q9ywU1cbCE/preview" width="740" height="420"></iframe>

____________

##  Syntax from math

The `babies` data set contains observations about the birthweight and other characteristics of children born in the San Francisco Bay area from 1960--1967.

We would like to build a model for birthweight as a function of the mother's age and whether this child was her first (`parity == 0`). Use the mathematical specification below to code the model in R.

$$birthweight=\beta_0 + \beta_1 \cdot age + \beta_2 \cdot parity + \varepsilon$$

__________

### Exercise{-}

The birthweight variable is recorded in the column `bwt`.

* Use `lm()` to build the parallel slopes model specified above. It's not necessary to use `factor()` in this case as the variable `parity` is coded using binary numeric values.

```{r}
# build model
lm(bwt ~ age + parity, data = babies)
```

__________

## Syntax from plot

This time, we'd like to build a model for birthweight as a function of the length of gestation and the mother's smoking status. Use Figure \@ref(fig:SmokP1) to inform your model specification.

```{r, label = "SmokP1", fig.cap = "`bwt` versus `gestation`", fig.width = 8}
ggplot(data = babies, aes(x = gestation, y = bwt, color = factor(smoke))) +
  geom_point(alpha = 0.5) +
  theme_bw()
```

_______

### Exercise{-}

* Use `lm()` to build a parallel slopes model implied by the plot. It's not necessary to use `factor()` in this case either.

```{r}
# build model
lm(bwt ~ gestation + smoke, data = babies)
```

```{r}
babies <- babies %>% 
  mutate(smoke = factor(smoke))
ggplot(data = babies, aes(x = gestation, y = bwt, color = smoke)) + 
         geom_point() + 
         geom_parallel_slopes(se = FALSE) + 
         theme_bw()
```


______________
