--- 
title: "[Machine Learning Toolbox](https://www.datacamp.com/courses/machine-learning-toolbox)"
author: "[Your Name Here](https://your_github_username.github.io/)"
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "The output format used for these personal notes is bookdown::gitbook."
---

# Prerequisites

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center", fig.width = 4, fig.height = 4)
library(ggplot2)
library(caret)
library(mlbench)
library(caTools)
# Parallel Processing
library(doMC)
registerDoMC(cores = 12)
```

This material is from the [DataCamp](https://www.datacamp.com) course [Machine Learning Toolbox](https://www.datacamp.com/courses/machine-learning-toolbox) by Zachary Deane-Mayer and Max Kuhn.  Before using this material, the reader should have completed and be comfortable with the material in the DataCamp modules [Introduction to R](https://www.datacamp.com/courses/free-introduction-to-r), [Intermediate R](https://www.datacamp.com/courses/intermediate-r), and [Correlation and Regression](https://www.datacamp.com/courses/correlation-and-regression).


Reminder to self: each `*.Rmd` file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.



```{r include=FALSE, eval = TRUE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'caret', 
  'caTools', 'ggplot2', "ranger", "randomForest"
), 'packages.bib')
```
